---
title: "Stat 154 EDA functions"
author: "Emily Liang"
date: "11/26/2017"
output: html_document
---
#Data we are working with
```{r}
library(readr)

df <- read.table("~/Documents/Academics/Stat 154/adult.data.txt", sep = ",",
                         colClasses=c("numeric", "character", "numeric", "factor",
                        "numeric", "factor", "character", "factor", "factor", "factor",
                        rep("numeric", 3),
                        "character","factor"))
names(data)<- c("age", "workclass", "final-weight", "education", "education-num", "marital_status", "occupation", "relationship", "race", "sex", "capital-gain", "capital-loss", "hours-per-week", "native_country", "income")

df <- as.data.frame(df)

#Test data
df_test <- read.table("~/Documents/Academics/Stat 154/adult.test.txt",skip =1, sep = ",", colClasses=c("numeric", "character", "numeric", "factor",
                        "numeric", "factor", "character", "factor", "factor", "factor",
                        rep("numeric", 3),
                        "character","factor"))
names(af_test)<- c("age", "workclass", "final-weight", "education", "education-num", "marital_status", "occupation", "relationship", "race", "sex", "capital-gain", "capital-loss", "hours-per-week", "native_country", "income")
df_test <- as.data.frame(df_test)
```

Libraries used:
```{r}
library(readr)
library(dplyr)
library(party)
library(rpart)
library(rpart.plot)
library(ROCR)
library(class)
library(Hmisc)
library(missForest)

set.seed(100) #not sure if this is necessary but just in case
```

###Missing/NA values:
```{r}
NA_imputes <- function(data){
  #Dealing with NAs
  data[data == ' ?'] <- NA #Change all '?' values to NA
  NA_Columns <-   colSums(is.na(data)) #per column
  NA_percent <- colMeans(is.na(data)*100) #percents
  
  #Convert back to factors
    data$workclass <- as.factor(data$workclass)
    data$occupation <- as.factor(data$occupation)
    data$`native_country` <- as.factor(data$`nativecountry`)
  
    #Imputation
      imputationResults <- missForest(xmis = data, maxiter = 2, variablewise = TRUE)
      dataMissForestImputed <- imputationResults$ximp #this is the adult data set with all the imputations
      #plots comparing missing values:
  dataMissForestImputed
}
```

###Plots of before and after imputation
```{r}
plot(df$occupation)
plot(dataMissForestImputed$occupation, col = rgb(1,0,0,1/4), add=T)


plot(df$workclass, col = rgb(0, 1, 1, 1/8))
plot(dataMissForestImputed$workclass, col = rgb(1,0,0,1/10), add=T)

plot(df$native_country, col = rgb(0, 1, 1, 1/8))
plot(dataMissForestImputed$native_country, col = rgb(1,0,0,1/16), add=T)
```

Given that we have now imputed our missing values correctly, we can move onto examining our various features of the data.

//

##Cleaning the variables

###Occupation
Splitting up Occupation was largely based on three levels - white vs blue collar, service vs not service, and then skilled vs. not skilled for not service jobs.
  
```{r} 
occupation_combine <- function(data){
    data$occupation =   
          gsub("^ Adm-clerical","White Collar Not Skilled",data$occupation)
    data$occupation =   
          gsub("^ Armed-Forces","Military",data$occupation)
    data$occupation =   
          gsub("^ Craft-repair","Blue-Collar Services",data$occupation)
    data$occupation = 
          gsub("^ Exec-managerial","White-Collar Skilled",data$occupation)
    data$occupation = 
          gsub("^ Farming-fishing","Blue-Collar Skilled",data$occupation)
    data$occupation = 
          gsub("^ Handlers-cleaners","Other Services",data$occupation)
    data$occupation = 
          gsub("^ Machine-op-inspct","Blue-Collar Skilled",data$occupation)
    data$occupation = 
          gsub("^ Other-service","Other Services",data$occupation)
    data$occupation = 
          gsub("^ Priv-house-serv","Other Services",data$occupation)
    data$occupation = 
          gsub("^ Prof-specialty","White-Collar Skilled",data$occupation)
    data$occupation = 
          gsub("^ Protective-serv","White Collar Services",data$occupation)
    data$occupation = 
          gsub("^ Sales","White Collar Services",data$occupation)
    data$occupation =   
          gsub("^ Tech-support","White Collar Services",data$occupation)
    data$occupation = 
          gsub("^ Transport-moving","Blue-Collar Services",data$occupation)
} 
```

###Native_country
Splitting up countries was largely based on geographic location (namely segments of Latin America, Europe, and Asia), political relations, and the level of development of the economy (using the HDI index for 1994, as well as examining the list of developed and developing countries). Countries with strong political ties, close geographic location, and similar economies tended to be grouped together. There are at least two groups per continent (of the countries within the data). 

The groupings are: 
  
  * Carribean-1: Columbia, Dominican-Republic, Outlying-US(Guam-USVI-etc)
  * Carribean-2: Jamaica, Puerto Rico, Trinadad&Tobago, Haiti
  * Canada: Canada
  * US: United States, Cuba
  * Central-America: El-Salvador, Guatemala, Honduras, Mexico, Nicaragua
  * Central-Europe: Germany, Greece, Italy, Yugoslavia
  * Central-West: France, Holand-netherlands
  * China: China, Hong
  * East-Europe: Hungary, Poland, South
  * England: England
  * Ireland-Scotland: Ireland, Scotland
  * Japan: Japan
  * South America: Ecuador, Peru
  * South Asia, 1: India, Iran, Cambodia
  * South Asia, 2: Laos, Thailand, Vietnam 
  * South Asia, 3: Philippines, Taiwan
  
```{r}
country_combine <- function(data){
      data$native_country = gsub(
          "^ Cambodia","South Asia, 1",data$native_country)
      data$native_country = gsub(
          "^ Canada","Canada",data$native_country)
      data$native_country = gsub(
          "^ China","China",data$native_country)
       data$native_country = gsub(
          "^ Columbia","Carribean-1",
          data$native_country)
      data$native_country = gsub(
          "^ Cuba","US",data$native_country)
      data$native_country = gsub(
          "^ Dominican-Republic","Carribean-1",
          data$native_country)
      data$native_country = gsub(
          "^ Ecuador","South America",data$native_country)
      data$native_country = gsub(
          "^ El-Salvador","Central-America",
          data$native_country)
      data$native_country = gsub(
          "^ England","England",data$native_country)
      data$native_country = gsub(
          "^ France","Central-West",data$native_country)
      data$native_country = gsub(
          "^ Germany","Central-Europe",data$native_country)
      data$native_country = gsub(
          "^ Greece","Central-Europe",data$native_country)
      data$native_country = gsub(
          "^ Guatemala","Central-America",
          data$native_country)
      data$native_country = gsub(
          "^ Haiti","Carribean-2",data$native_country)
      data$native_country = gsub(
          "^ Holand-Netherlands","Central-West",
          data$native_country)
      data$native_country = gsub(
          "^ Honduras","Central-America",
          data$native_country)
      data$native_country = gsub(
          "^ Hong","China",data$native_country)
      data$native_country = gsub(
          "^ Hungary","East-Europe",data$native_country)
      data$native_country = gsub(
          "^ India","South Asia, 1",data$native_country)
      data$native_country = gsub(
          "^ Iran","South Asia, 1",data$native_country)
      data$native_country = gsub(
          "^ Ireland","Ireland-Scotland",
          data$native_country)
      data$native_country = gsub(
          "^ Italy","Central-Europe",data$native_country)
      data$native_country = gsub(
          "^ Jamaica","Carribean-2",data$native_country)
      data$native_country = gsub(
          "^ Japan","Japan",data$native_country)
      data$native_country = gsub(
          "^ Laos","South Asia, 2",data$native_country)
      data$native_country = gsub(
          "^ Mexico","Central-America",
          data$native_country)
      data$native_country = gsub(
          "^ Nicaragua","Central-America",
          data$native_country)
      data$native_country = gsub(
          "^ Peru","South America",data$native_country)
      data$native_country = gsub(
          "^ Philippines","South Asia, 3",
          data$native_country)
      data$native_country = gsub(
          "^ Poland","East-Europe",data$native_country)
      data$native_country = gsub(
          "^ Portugal","West-Europe",data$native_country)
      data$native_country = gsub(
          "^ Puerto-Rico","Carribean-2",
          data$native_country)
      data$native_country = gsub(
          "^ Scotland","Ireland-Scotland",
          data$native_country)
      data$native_country = gsub(
          "^ South","East-Europe",data$native_country)
      data$native_country = gsub(
          "^ Taiwan","South Asia, 3",data$native_country)
      data$native_country = gsub(
          "^ Thailand","South Asia, 2",data$native_country)
      data$native_country = gsub(
          "^ Trinadad&Tobago","Carribean-2",
          data$native_country)
      data$native_country = gsub(
          "^ United-States","US",data$native_country)
      data$native_country = gsub(
          "^ Vietnam","South Asia, 2",data$native_country)
      data$native_country = gsub(
          "^ Yugoslavia","Central-Europe",data$native_country)
      #Having some issues with this one:
      vecc <- which(data$native_country == 
                      " Outlying-US(Guam-USVI-etc)")
      data$native_country[vecc] <- "Carribean-1"
}
      
tbl_country <- table(data$native_country, data$income) 
tbl_country
```

###Workclass
The workclass variable currently has 8 different levels - Federal-gov, Local-goc, State-gov, Private, Self-emp-inc, Self-emp-not-inc, Without pay, and Never-worked. We decided to group all the government jobs together, after examining the income ratio for each. We did not combine the self employed classes together, given that self employed incorporated positions tend to have higher pay than not-incorporated ones. Finally, We combined those that never worked with those that were without pay into a new group, of those who were not working. 

```{r}
workclass_combine <- function(data){
  data$workclass = gsub(
        "^ Federal-gov","Gov",data$workclass)
    data$workclass = gsub(
        "^ Local-gov","Gov",data$workclass)
    data$workclass = gsub(
        "^ State-gov","Gov",data$workclass)
    data$workclass = gsub(
        "^ Private","Private",data$workclass)
    data$workclass = gsub(
        "^ Self-emp-inc","Self-Employed-Incorp",data$workclass)
    data$workclass = gsub(
        "^ Self-emp-not-inc","Self-Employed-Not-Incorp",data$workclass)
    data$workclass = gsub(
        "^ Without-pay","Not-Working",data$workclass)
    data$workclass = gsub(
        "^ Never-worked","Not-Working",data$workclass)
}

tbl_work <- table(data$workclass, data$income) 

```

###Marital_status
For marital status, we split the groups into based on whether or not that person was married, considered to legally have a spouse, and whether or not their spouse was living with them. "Has-Spouse" indicates that the person was married and with their spouse (Married-civ-spouse for civilian spouses, and Married-AF-spouse for spouses from the armed forces), whereas "Married-spouse-absent" and "Separated" denoted an "Absent-Spouse". Divorced and Widowed were placed under the same category as "No-Spouse", while Never-married stayed the same but was renamed "Single". 

```{r}
marital_combine <- function(data){
  data$martial_status = gsub(
        "^ Divorced","No-Spouse",data$martial_status)
    data$martial_status = gsub(
        "^ Widowed","No-Spouse",data$martial_status)
    data$martial_status = gsub(
        "^ Married-spouse-absent","Absent-Spouse",data$martial_status)
    data$martial_status = gsub(
        "^ Never-married","Single",data$martial_status)
    data$martial_status = gsub(
        "^ Separated","Absent-Spouse",data$martial_status)
    
    data$martial_status = gsub(
        "^ Married-AF-spouse","Has-Spouse",data$martial_status)
    data$martial_status = gsub(
        "^ Married-civ-spouse","Has-Spouse",data$martial_status)
}

tbl_marital <- table(data$martial_status, data$income)
tbl_marital

```


###Relationships
For relationships, this is the relationship of the surveyor to a central household figure. Unmarried referred to an unmarried partner; as such, we combined that we "Not-in-family" to make a grouping of Not-relative relationships. Husband and Wife were grouped together under "Spouse", while the last two remaining classes were combined as "Other-relative" relationships.

```{r}
relation_combine <- function(data){
  data$relationship = gsub(
        "^ Husband","Spouse",data$relationship)
    data$relationship = gsub(
        "^ Wife","Spouse",data$relationship)
    data$relationship = gsub(
        "^ Other-relative","Other-relative",data$relationship)
    data$relationship = gsub(
        "^ Own-child","Other-relative",data$relationship)
    data$relationship = gsub(
        "^ Unmarried","Not-relative",data$relationship)
    data$relationship = gsub(
        "^ Not-in-family","Not-relative",data$relationship)
}

tbl_relat <- table(data$relationship, data$income)
tbl_relat
```

#Income
For income (1 if over 50K, 0 if less)
```{r}
Income_factor <- function(data){
  data$income = gsub(
    "^ <=50K",0,data$income)
  data$income = gsub(
    "^ >50K",1,data$income)
}
```

###Education
Education:
As education and education-num represented essentially the same data, we decided to remove one of them from our dataset. Education-num could also be slightly misleading, since the "10th grade" is labeled as having "5 years" of education, and so forth. As such, we decided it would be better to keep education and focus on that feature.
```{r}
data$`education-num` <- NULL
```

###Capital_gain
Capital_gain:
We noticed that there appear to be outliers for capital gain (there was a chunk of data that had values of over 99999). As such, we decided to change these outlier values to the mean value of the capital gains.
```{r}
cap_gain_clean_outlier <- function(data){
  mean_cg <- mean(data$'capital-gain')
  data$`capital-gain`[which(data$`capital-gain` == 99999)] <- mean_cg
} 
```

###Hours
For the hours-per-week worked, we decided to bin the data together based off of their quartiles. We examined where the cut-offs of each percentile (of 10%) was. Based off of those unique cut-off points, we generated 6 levels to bin our data to:

- [1,24]
- (24,35]
- (35,40]
- (40,48]
- (48,55]
- (55,99]

```{r}
quant <- quantile(df$`hours-per-week`,seq(0,1,.10))

hours_combine <- function(data){
  data$`hours-per-week` <-
  cut(data$`hours-per-week`,
      unique(quant),include.lowest=TRUE)
}
```

Here is a function to make sure everything is a factor by the end:
```{r}
factorize <- function(data){
    cols <- c("age", "workclass", 'education', 'martial_status', 'occupation', 'relationship', 'race', 'sex', 'capital-gain', 'capital-loss', 'native_country', 'income') #no finl-weight because that is continuous
    data[cols] <- lapply(data[cols], as.factor)
}
```


```{r}
age_median = function(dat){
  Age_med_inc = c(19340, 33151, 41667, 47261, 35232,  21422, 14731)
  xf <- as.factor(dat$age)
  dat$age <- Age_med_inc[xf]
  dat
}

```

###Make occupation into numerical variable
```{r}
occup_sex_med = function(dat){
  interacts <- interaction(dat$sex,dat$occupation)
  medians <- c(21023, 41410, 10905, 26449, 25980, 7059, 41090, 15880, 25790, 24681, 21546, 10431, 20621, 22524, 30642, 11823, 11746,9070, 15815, 12095, 11686, 7453, 7893, 23656, 3287, 4223, 11746)
  x.lvl <- unique(interacts)
    dat$occ_sex <- medians[match(interacts, x.lvl)]
    dat
}
```
